Manuscript

### 2015-01-26 inst/simulation/manuscript.Rmd
...

  #plot(tracksSP, col="grey", xlim=c(3.4, 3.5)*1e6, ylim=c(7.0, 7.1)*1e6, asp=1)

  intersections$tracks$tracks <- subset(intersections$tracks$tracks, year==2001)
  y <- intersections$tracks$toGGDF()
  y$scenario <- scenario
  y$observation <- T
  levels(y$group) <- paste(levels(y$group), "2")
  trackData <- rbind(trackData, x, y)
  #plot(tracksSP, col="black", add=T)

  x <- intersections$surveyRoutes$toGGDF()
  x$scenario <- scenario    
  observedAt <- intersections$intersections$intersections[intersections$intersections$intersections$intersections > 0 & intersections$intersections$intersections$year == 2001,]$surveyRoute
  y <- data.frame(id=unique(x$id), observed=FALSE)
  if (length(observedAt) > 0) y[y$id %in% observedAt,]$observed <- TRUE
  x <- plyr::join(x, y, by="id")
  surveyRoutesData <- rbind(surveyRoutesData, x)

  #plot(intersections$surveyRoutes$surveyRoutes, col="blue", add=T)    
  i <- i + 1
}

@@

```{r figure-2-init}
getSurveyRouteLocations = function(scenarios, modelName, iterations, years) {
  ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, years) {
    message("Scenario = ", x$scenario, ", iteration = ", x$iteration)
    study <- getStudy(scenario=x$scenario)
    estimates <- study$getModel(modelName=modelName, iteration=x$iteration) 
    estimates$modelName <- modelName
    estimates <- study$loadEstimates(estimates)
    intersections <- subset(estimates$data, year %in% years)$intersections>0
    surveyRoutes <- data.frame(estimates$getUnscaledObservationCoordinates(), Intersection=intersections)
    surveyRoutes$Scenario <- x$scenario
    surveyRoutes$Iteration <- x$iteration
    return(surveyRoutes)
  }, modelName=modelName, years=years)
}

getAgentLocations = function(scenarios, modelName, iterations, years) {
  ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, years) {
    study <- getStudy(scenario=x$scenario)
    tracks <- study$loadTracks(iteration=x$iteration)
    tracks$tracks <- subset(tracks$tracks, year %in% years & yday<59)
    intersections <- study$loadIntersections(iteration=x$iteration)

    # If no intersection, randomized track can be anywhere else except on the survey route

    observedBursts <- aaply(intersections$intersectionsMatrix, 2, function(x) if (any(x>0)) TRUE else FALSE)
    observationTracks <- tracks$randomizeObservationDayTracks(days=1)

    a <- ddply(observationTracks$tracks, .(id, year, burst), function(x, observationTracks) {
      name <- with(x[1,], paste(id, year, burst))
      if (observedBursts[name] == FALSE) return(data.frame(x=mean(x$x), y=mean(x$y), year=x$year))
      NULL
    }, observationTracks=observationTracks)[,c("x","y","year")]
    a$Observed <- F

    index <- intersections$intersections$intersections > 0 & intersections$intersections$year %in% years
    if (all(index == FALSE)) b <- data.frame() else {
      x <- intersections$intersections[index,]
      b <- data.frame(coordinates(x), year=x$year)
      b$Observed <- T
    }
    agentLocations <- rbind(a, b)
    agentLocations$Scenario <- x$scenario
    agentLocations$Model <- x$modelName

@@

getEstimatedDensity = function(scenarios, modelName, iterations, year, sd=FALSE, log=FALSE) {
  ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, year0, sd) {
    message("Scenario = ", x$scenario, ", iteration = ", x$iteration)
    study <- getStudy(scenario=x$scenario)
    estimates <- study$getModel(modelName=modelName, iteration=x$iteration)
    estimates$modelName <- modelName
    estimates <- study$loadEstimates(estimates)
    estimates$collectEstimates()

@@

    estimates$data <- subset(estimates$data, year == year0)
    if (log) estimates$data$fittedMean <- log(estimates$data$fittedMean)
    density <- estimates$getPopulationDensityInterpolate(getSD=sd, maskPolygon=studyArea$boundary)
    density <- if (sd) density$sd else density$mean
    if (is.null(density) || maxValue(density$rasterStack[[1]]) > 1) stop("Estimation failed.")
    density <- as.data.frame(density$rasterStack[[1]], xy=T)
    density$Scenario <- x$scenario
    density$Iteration <- x$iteration
    density$Model <- modelName
    density$Year <- year0

    return(density)
  }, modelName=modelName, year0=year, sd=sd)
}
```

```{r figure-2-test}
iterations <- as.integer(c(1,1,3,2,5,6))
year <- 2010
densityST <- getEstimatedDensity(scenarios[1], modelNames[1], iterations[1], year=year, sd=FALSE)
p <- ggplot(densityST) + geom_raster(aes_string(x="x", y="y", fill=paste0("X", year))) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") 
saveFigure(p, "tmp_compare_densities_ST.png", width=8)
###

iterations <- as.integer(c(1,1,3,2,5,6))
year <- 2010
surveyRoutes <- getSurveyRouteLocations(scenarios=scenarios[1], modelName=modelNames[1], iterations=iterations[1], years=year)
surveyRoutes <- rbind(surveyRoutes[surveyRoutes$Intersection == FALSE,], surveyRoutes[surveyRoutes$Intersection == TRUE,])
agents <- getAgentLocations(scenarios=scenarios[1], modelName=modelNames[1], iterations=iterations[1], years=year)
densityST <- getEstimatedDensity(scenarios[1], modelNames[1], iterations[1], year=year, sd=FALSE)
densityFMP <- getEstimatedDensity(scenarios[1], modelNames[3], iterations[1], year=year, sd=FALSE)
limits <- extent(study$studyArea$boundary)

p <- ggplot(densityST) + geom_raster(aes_string(x="x", y="y", fill=paste0("X", year))) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") +
  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black")) + 
  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal()
print(p)
saveFigure(p, "tmp_compare_densities_ST.png", width=8)

p <- ggplot(densityFMP) + geom_raster(aes_string(x="x", y="y", fill=paste0("X", year))) + scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent") +
  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black")) + 
  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal()
print(p)
saveFigure(p, "tmp_compare_densities_FMP.png", width=8)
```

```{r figure-2}
#iterations <- as.integer(c(1,1,3,2,5,6))
##iterations <- as.integer(c(1,1,3,2,5,7))
iterations <- as.integer(c(1,1,3,2,5,8))
year <- 2010
yearLayer <- paste0("X", year)
surveyRoutes <- getSurveyRouteLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
surveyRoutes <- rbind(surveyRoutes[surveyRoutes$Intersection == FALSE,], surveyRoutes[surveyRoutes$Intersection == TRUE,])
agents <- getAgentLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, sd=F, log=T)
# TODO: do not use scaling??
#densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer) { x[,yearLayer] <- scale(x[,yearLayer]); return(x) }, yearLayer=yearLayer)
study <- getStudy(scenario="A")
limits <- extent(study$studyArea$boundary)

# Scale density to 1/km^2
densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer, weight) {
  x[,yearLayer] <- x[,yearLayer] * weight
  return(x)
}, yearLayer=yearLayer, weight=1 / (prod(res(study$getTemplateRaster())) / 1000^2))

#p <- ggplot(boundaryDF) + geom_polygon(aes(long, lat, group=group), fill=terrain.colors(100)[1]) +
p <- ggplot(densitiesScaled) + geom_raster(aes_string(x="x", y="y", fill=yearLayer)) +
  scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent", guide=guide_legend(title=expression(paste("Population density ", (individuals/km^2))))) +
  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
  geom_point(data=agents, aes(x, y, colour=Observed), size=1.3, colour="white") +
  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black"), guide=FALSE) +
  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
  facet_grid(~scenario) + coord_equal() + theme_raster(base_size=10) + theme(legend.position="bottom")
print(p)
saveFigure(p, "manuscript-figure-2a.png", width=8)

@@

p <- ggplot(boundaryDF) + geom_path(aes(long, lat, group=group), colour="grey") +
  geom_point(data=surveyRoutes, aes(x, y, colour=Intersection), size=1) + scale_color_manual(values=c("darkgrey", "black")) +
  scale_x_continuous(breaks=seq(gridTemplate@extent@xmin, gridTemplate@extent@xmax, 1e5), expand=c(0,0)) +
  scale_y_continuous(breaks=seq(gridTemplate@extent@ymin, gridTemplate@extent@ymax, 1e5), expand=c(0,0)) +
  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal() + theme(panel.grid.major=element_line(colour="grey"))
print(p)
saveFigure(p, "manuscript-figure-2b.png", width=8)
```

>>>>

  intersections$tracks$tracks <- subset(intersections$tracks$tracks, year==2001)
  y <- intersections$tracks$toGGDF()
  y$scenario <- scenario
  y$observation <- T
  levels(y$group) <- paste(levels(y$group), "2")
  trackData <- rbind(trackData, x, y)

  x <- intersections$surveyRoutes$toGGDF()
  x$scenario <- scenario    
  observedAt <- intersections$intersections$intersections[intersections$intersections$intersections$intersections > 0 & intersections$intersections$intersections$year == 2001,]$surveyRoute
  y <- data.frame(id=unique(x$id), observation=FALSE)
  if (length(observedAt) > 0) y[y$id %in% observedAt,]$observation <- TRUE
  x <- plyr::join(x, y, by="id")
  surveyRoutesData <- rbind(surveyRoutesData, x)

  #trackData$type <- "track"
  #surveyRoutesData$type <- "transect"
  #x <- rbind(trackData[,c("long","lat","order","piece","group","observation","type")], surveyRoutesData[,c("long","lat","order","piece","group","observation","type")])
  #x$colour <- factor(paste(x$type, x$observation))
  #ggplot(x, aes(long, lat, group=group, colour=colour)) + geom_path()

  i <- i + 1
}

@@

```{r, eval=F}
#x<-data.frame(scenario="C", iteration=as.integer(3), stringsAsFactors=F)
modelName <- modelNames[2]
x<-data.frame(scenario="A", iteration=as.integer(10), stringsAsFactors=F)
study <- getStudy(scenario=x$scenario)
estimates <- study$getModel(modelName=modelName, iteration=x$iteration)
estimates$modelName <- modelName
estimates <- study$loadEstimates(estimates)
intersections <- study$loadIntersections(iteration=x$iteration)
head(subset(intersections$intersections@data, intersections>0), 20)
head(subset(estimates$data, intersections>0), 20)

model <- SimulatedSmoothModelTemporal(study=study, iteration=as.integer(1))
params <- list(family="nbinomial", offsetScale=1000^2, timeModel="ar1")
intersections <- study$loadIntersections(iteration = model$iteration)
model$setup(intersections = intersections, params = params)
model$estimate()

head(subset(model$data, intersections>0), 20)
head(subset(intersections$intersections@data, intersections>0), 20)

./count_intersections.sh ~/git/RParallelScreen/ A 1 1 1
./estimate.sh ~/git/RParallelScreen/ A 1 1 SmoothModel-nbinomial-ar1

./estimate.sh ~/git/RParallelScreen/ A 1 1 FMPModel

```

```{r figure-2-init}
getSurveyRouteLocations = function(scenarios, modelName, iterations, years) {
  ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, years) {
    message("Scenario = ", x$scenario, ", iteration = ", x$iteration)
    study <- getStudy(scenario=x$scenario)
    estimates <- study$getModel(modelName=modelName, iteration=x$iteration) 
    estimates$modelName <- modelName
    estimates <- study$loadEstimates(estimates)
    z <- cbind(estimates$getUnscaledObservationCoordinates(), estimates$data)
    surveyRoutes <- subset(z, year %in% years, select=c(x,y,intersections))
    surveyRoutes$intersections <- surveyRoutes$intersections > 0
    surveyRoutes$Scenario <- x$scenario
    surveyRoutes$Iteration <- x$iteration
    return(surveyRoutes)
  }, modelName=modelName, years=years)
}

#x<-data.frame(scenario="C", iteration=as.integer(3), stringsAsFactors=F)
getAgentLocations = function(scenarios, modelName, iterations, years) {
  ddply(data.frame(scenario=scenarios, iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, years) {

    #study <- getStudy(scenario=x$scenario)
    #intersections <- study$loadIntersections(iteration=x$iteration)
    #observationTracks <- SimulatedTracks$new(study=study, iteration=x$iteration)
    #observationTracks$loadTracks(fileName=observationTracks$getObservationTracksFileName(), addColumns=TRUE)
    # TODO: use observation tracks ^^^^

    study <- getStudy(scenario=x$scenario)
    tracks <- study$loadTracks(iteration=x$iteration)
    tracks$tracks <- subset(tracks$tracks, year %in% years & yday<59)
    intersections <- study$loadIntersections(iteration=x$iteration)

    # The observation day has not been recorded, ouch... just randomize a day from the 60 days track
    observedBursts <- aaply(intersections$intersectionsMatrix, 2, function(x) if (any(x>0)) TRUE else FALSE)
    #observationTracks <- tracks$randomizeObservationDayTracks(days=1)
    observationTracks <- tracks

    #d_ply(observationTracks$tracks, .(id, year, burst), function(x) {
    #  with(x[1,], message(id, " ", year, " ", burst))
    #})

    # Return mean location of the unobserved agents
    a <- ddply(observationTracks$tracks, .(id, year, burst), function(x, observedBursts) {
      name <- with(x[1,], paste(id, year, burst))
      if (observedBursts[name] == FALSE) return(data.frame(x=mean(x$x), y=mean(x$y), year=x$year[1]))
      #if (observedBursts[name] == FALSE) return(data.frame(x=x$x, y=x$y, year=x$year))
      NULL
    }, observedBursts=observedBursts)[,c("x","y","year")]
    a$Observed <- F

    # But for the observed agents we know the location, which is the transect
    index <- intersections$intersections$intersections > 0 & intersections$intersections$year %in% years
    b <- if (all(index == FALSE)) data.frame() else {
      x <- intersections$intersections[index,]
      y <- data.frame(coordinates(x), year=x$year)
      y$Observed <- T
      y
    }

    agentLocations <- rbind(a, b)
    agentLocations$Scenario <- x$scenario
    agentLocations$Model <- x$modelName

@@

getEstimatedDensity = function(scenarios, modelName, iterations, year, sd=FALSE, log=FALSE) {
  ddply(data.frame(scenario=scenarios, year=rep(year, scenarios), iteration=iterations, stringsAsFactors=FALSE), .(scenario, iteration), function(x, modelName, sd) {
    message("Scenario = ", x$scenario, ", iteration = ", x$iteration)
    study <- getStudy(scenario=x$scenario)
    estimates <- study$getModel(modelName=modelName, iteration=x$iteration)
    estimates$modelName <- modelName
    estimates <- study$loadEstimates(estimates)
    estimates$collectEstimates()

@@

    estimates$data <- subset(estimates$data, year == x$year)
    #if (log) estimates$data$fittedMean <- log(estimates$data$fittedMean)
    density <- estimates$getPopulationDensityInterpolate(getSD=sd, maskPolygon=studyArea$boundary)
    density <- if (sd) density$sd else density$mean
    if (is.null(density) || maxValue(density$rasterStack[[1]]) > 1) stop("Estimation failed.")
    density <- as.data.frame(density$rasterStack[[1]], xy=T)
    density$Scenario <- x$scenario
    density$Iteration <- x$iteration
    density$Model <- modelName
    density$Year <- year0

    return(density)
  }, modelName=modelName, sd=sd)
}
```

```{r figure-2}
study <- getStudy(scenario="A")
limits <- extent(study$studyArea$boundary)
iterations <- as.integer(c(1,6,3,2,5,31))
densities <- getEstimatedDensity(scenarios=scenarios[6], modelName=modelNames[1], iterations=iterations[6], year=year, sd=F, log=F)
year <- 2010
yearLayer <- paste0("X", year)
surveyRoutes <- getSurveyRouteLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
#surveyRoutes <- rbind(surveyRoutes[surveyRoutes$intersections == FALSE,], surveyRoutes[surveyRoutes$intersections == TRUE,])
agents <- getAgentLocations(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, years=year)
densities <- getEstimatedDensity(scenarios=scenarios, modelName=modelNames[1], iterations=iterations, year=year, sd=F, log=F)
# TODO: do not use scaling??
#densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer) { x[,yearLayer] <- scale(x[,yearLayer]); return(x) }, yearLayer=yearLayer)

# Scale density to 1/km^2
densitiesScaled <- ddply(densities, .(scenario), function(x, yearLayer, weight, logTrans) {
  x[,yearLayer] <- x[,yearLayer] * weight
  if (logTrans) x[,yearLayer] <- log(x[,yearLayer])
  return(x)
}, yearLayer=yearLayer, weight=1 / (prod(res(study$getTemplateRaster())) / 1000^2), logTrans=F)

p <- ggplot(densitiesScaled) + geom_raster(aes_string(x="x", y="y", fill=yearLayer)) +
  scale_fill_gradientn(colours=terrain.colors(100), na.value="transparent", guide=guide_legend(title=expression(paste("Population density ", (individuals/km^2))))) +
  geom_path(data=boundaryDF, aes(long, lat, group=group), colour="darkgreen") +
  geom_point(data=agents, aes(x, y, colour=Observed), size=1.3, colour="white") +
  geom_point(data=agents, aes(x, y, colour=Observed), size=1) + scale_color_manual(values=c("darkgray", "black"), guide=FALSE) +
  scale_x_continuous(limits=c(limits@xmin, limits@xmax), expand=c(0,0)) +
  scale_y_continuous(limits=c(limits@ymin, limits@ymax), expand=c(0,0)) +
  facet_grid(~scenario) + coord_equal() + theme_raster(base_size=10) + theme(legend.position="bottom")
print(p)
saveFigure(p, "manuscript-figure-2a.png", width=6,height=3)

@@

p <- ggplot(boundaryDF) + geom_path(aes(long, lat, group=group), colour="grey") +
  geom_point(data=surveyRoutes, aes(x, y, colour=intersections), size=1) + scale_color_manual(values=c("darkgrey", "black")) +
  scale_x_continuous(breaks=seq(gridTemplate@extent@xmin, gridTemplate@extent@xmax, 1e5), expand=c(0,0)) +
  scale_y_continuous(breaks=seq(gridTemplate@extent@ymin, gridTemplate@extent@ymax, 1e5), expand=c(0,0)) +
  facet_grid(~scenario) + theme_raster(base_size=10) + coord_equal() + theme(panel.grid.major=element_line(colour="grey"))
print(p)
saveFigure(p, "manuscript-figure-2b.png", width=6,height=3)
```
